---
title: "Reproducible Research_Course Project 1"
output:
  html_document:
    df_print: paged
---

## Loading and preprocessing the data


```{r}
activity <- read.csv("C:/Users/JJQ/Documents/data/activity.csv", header = TRUE, stringsAsFactors = FALSE)
summary(activity)
```

## What is mean total number of steps taken per day?

```{r}
library(dplyr)

steps <- activity %>% select(steps, date) %>% group_by(date) %>% summarise(steps_per_day = sum(steps))

hist(steps$steps_per_day, main = "Total Number of Steps Taken per Day", xlab = "Steps Per Day", ylab = "Interval", col = "light blue")

mean_dailysteps <- mean(steps$steps_per_day, na.rm = TRUE)
mean_dailysteps

median_dailysteps <- median(steps$steps_per_day, na.rm = TRUE)
median_dailysteps
```

Answer: 
The mean of the total number of steps taken per day is equal to 10766.19
The median of the total number of steps taken per day is equal to 10765

## What is the average daily activity pattern?
```{r}
library(ggplot2)

average <- group_by(activity, interval) %>% summarise(average_steps = mean(steps, na.rm = TRUE))

ggplot(average, aes(x = interval, y = average_steps, type = "l")) + geom_line(color = "blue") + 
  ggtitle("Average Number of Steps by Interval") + xlab("Interval") + ylab("Average Steps across All Days")
  
interval_max_steps <- filter(average, average_steps == max(average$average_steps))
interval_max_steps
```

Answer: interval 835 contains the maximum number of steps. The person is very active at around 2:00pm (to be accurate at 1:55pm)

## Imputig missing values
```{r}
summary(is.na(activity))

activity[is.na(activity)] <- c(average$average_steps)
activity_new <- activity

steps_new <- activity_new %>% select(steps, date) %>% group_by(date) %>% summarise(steps_per_day_new = sum(steps))

hist(steps_new$steps_per_day_new, main = "Total Number of Steps Taken per Day (New Dataset)", xlab = "Steps Per Day (New)", ylab = "Interval", col = "blue")

mean_dailysteps_new <- mean(steps_new$steps_per_day_new, na.rm = TRUE)
mean_dailysteps_new

median_dailysteps_new <- median(steps_new$steps_per_day_new, na.rm = TRUE)
median_dailysteps_new
```

Answer: 
There are 2304 missing value in total in the dataset
The mean of the total number of steps taken per day didn't change after imputing, still equal to 10766.19
The median of the total number of steps taken per day increased slightly to 10766.19 from 10765, after imputing

## Are there differences in activity patterns between weekdays and weekends
```{r}
wkd <- weekdays(as.Date(strptime(activity_new$date, format =  "%Y-%m-%d")))
wkd <- as.factor(wkd)

activity_wkd <- cbind(activity_new, wkd)

levels(activity_wkd$wkd) <- list(weekday = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), 
                                     weekend = c("Saturday", "Sunday"))
levels(activity_wkd$wkd)

par(mfrow = c(2,1))
with(activity_wkd[activity_wkde$wkd == "weekday", ], plot(aggregate(steps ~ interval, FUN = mean), type = "l", main = "Activity in Weekdays"))
with(activity_wkd[activity_wkd$wkd == "weekend", ], plot(aggregate(steps ~ interval, FUN = mean), type = "l", main = "Activity in Weekends"))
```

Answer:
The person is more activate in weekends than in weekdays




